Supplementary Methods

### 2020-05-26 methods.tex
Remove URLs from references

% https://mail.python.org/pipermail/numpy-discussion/2015-October/073849.html
% https://github.com/numpy/numpy/pull/6352
NumPy adopted an official Governance Document on October~5,
2015\cite{NumPyProjectGovernance}.
Project decisions are usually made by consensus of interested contributors.
This means that, for most decisions, everyone is entrusted with veto power.
A Steering Council, currently composed of 12~members, facilitates this
process and oversees daily development of the project by contributing code
and reviewing contributions from the community.

% https://mail.python.org/pipermail/numpy-discussion/2018-July/078476.html
% https://github.com/numpy/numpy/pull/11865
NumPy's official Code of Conduct was approved on September~1, 2018\cite{NumPyCodeofConduct}.
In brief, we strive to:
\emph{be open};
\emph{be empathetic, welcoming, friendly, and patient};
\emph{be collaborative};
\emph{be inquisitive}; and
\emph{be careful in the words that we choose}.
The Code of Conduct also specifies how breaches can be reported and outlines
the process for responding to such reports.

@@

A Python \emph{wheel}\cite{PEP427} is a standard file format for
distributing Python libraries.  In addition to Python code, a
wheel may include compiled C extensions and other binary data.
This is important, because many libraries, including NumPy,
require a C compiler and other build tools to build the software
from the source code, making it difficult for many users to install
the software on their own.  The introduction of wheels to the Python
packaging system has made it much easier for users to install
precompiled libraries.

@@

The technology that is used to build the wheels evolves continually.
At the time this paper is being written, a key component is the
\texttt{multibuild} suite of tools developed by Matthew Brett and
other developers\cite{multibuild}.  Currently, scripts using
\texttt{multibuild} are written for the continuous integration
platforms Travis-CI (for Linux and Mac OSX) and Appveyor
(for Windows).

@@

To address these issues NumPy has the goal of providing the fundamental
API for \emph{interoperability} between the various NumPy-like APIs.
An earlier step in this direction was the implementation of the
\texttt{\_\_array\_ufunc\_\_} protocol in NumPy 1.13, which enabled interoperability
for most mathematical functions.\cite{NEP13}
In 2019 this was expanded more generally with the inclusion of the
\texttt{\_\_array\_function\_\_} protocol into NumPy~1.17.
These two protocols allow providers of array objects to be interoperable
with the NumPy API: their arrays work correctly with almost all NumPy functions.\cite{NEP18}
For the users relying on specialized array projects it means that even though
much code is written specifically for NumPy arrays and uses the NumPy API as
\texttt{import numpy as np}, it can nevertheless work for them.
For example, here is how a CuPy GPU array can be passed through NumPy for
processing, with all operations being dispatched back to CuPy:

@@

At the time of writing the two fastest supercomputers in the
world, Summit and Sierra, both have IBM POWER9 architectures
\cite{top500nov2019}. In late 2018, Astra, the first ARM-based
supercomputer to enter the TOP500 list, went into production\cite{
astra-wiki}. Furthermore, over 100 billion ARM processors have been
produced as of 2017\cite{arm-architecture}, making it the most 
widely used instruction set architecture in the world.

Clearly there are motivations for a large scientific computing
software library to support POWER and ARM architectures. We've extended
our continuous integration (CI) testing to include \texttt{ppc64le}
(POWER8 on Travis CI) and ARMv8 (on Shippable service). We also test
with the s390x architecture (IBM Z CPUs on Travis CI) so that we
can probe the behavior of our library on a big-endian machine.
This satisfies one of the major components of
improved CI testing laid out in a version of our roadmap
\cite{numpy-roadmap}---specifically, ``CI for more exotic
platforms."

PEP 599\cite{PEP599} lays out a plan for new Python binary wheel
distribution support, \texttt{manylinux2014}, that adds
support for a number of architectures supported by the CentOS
Alternative Architecture Special Interest Group, including
ARMv8, ppc64le, as well as s390x. We are thus well-positioned
for a future where provision of binaries on these architectures
will be expected for a library at the base of the ecosystem.

>>>>

% https://mail.python.org/pipermail/numpy-discussion/2015-October/073849.html
% https://github.com/numpy/numpy/pull/6352
NumPy adopted an official Governance Document on October~5,
2015 (\url{https://numpy.org/devdocs/dev/governance/governance.html}).
Project decisions are usually made by consensus of interested contributors.
This means that, for most decisions, everyone is entrusted with veto power.
A Steering Council, currently composed of 12~members, facilitates this
process and oversees daily development of the project by contributing code
and reviewing contributions from the community.

% https://mail.python.org/pipermail/numpy-discussion/2018-July/078476.html
% https://github.com/numpy/numpy/pull/11865
NumPy's official Code of Conduct was approved on September~1, 2018
(\url{https://numpy.org/devdocs/dev/conduct/code_of_conduct.html}).
In brief, we strive to:
\emph{be open};
\emph{be empathetic, welcoming, friendly, and patient};
\emph{be collaborative};
\emph{be inquisitive}; and
\emph{be careful in the words that we choose}.
The Code of Conduct also specifies how breaches can be reported and outlines
the process for responding to such reports.

@@

A Python \emph{wheel} ({https://www.python.org/dev/peps/pep-0427/})
is a standard file format for distributing Python libraries.
In addition to Python code, a wheel may include compiled
C extensions and other binary data.
This is important, because many libraries, including NumPy,
require a C compiler and other build tools to build the software
from the source code, making it difficult for many users to install
the software on their own.  The introduction of wheels to the Python
packaging system has made it much easier for users to install
precompiled libraries.

@@

The technology that is used to build the wheels evolves continually.
At the time this paper is being written, a key component is the
\texttt{multibuild} suite of tools developed by Matthew Brett and
other developers (\url{https://github.com/matthew-brett/multibuild}).
Currently, scripts using \texttt{multibuild} are written for
the continuous integration platforms Travis-CI (for Linux and Mac OSX)
and Appveyor (for Windows).

@@

To address these issues NumPy has the goal of providing the fundamental
API for \emph{interoperability} between the various NumPy-like APIs.
An earlier step in this direction was the implementation of the
\texttt{\_\_array\_ufunc\_\_} protocol in NumPy 1.13, which enabled interoperability
for most mathematical functions (\url{https://numpy.org/neps/nep-0013-ufunc-overrides.html}).
In 2019 this was expanded more generally with the inclusion of the
\texttt{\_\_array\_function\_\_} protocol into NumPy~1.17.
These two protocols allow providers of array objects to be interoperable
with the NumPy API: their arrays work correctly with almost all NumPy
functions (\url{https://numpy.org/neps/nep-0018-array-function-protocol.html}).
For the users relying on specialized array projects it means that even though
much code is written specifically for NumPy arrays and uses the NumPy API as
\texttt{import numpy as np}, it can nevertheless work for them.
For example, here is how a CuPy GPU array can be passed through NumPy for
processing, with all operations being dispatched back to CuPy:

@@

At the time of writing the two fastest supercomputers in the
world, Summit and Sierra, both have IBM POWER9 architectures
(\url{https://www.top500.org/lists/2019/11/}).
In late 2018, Astra, the first ARM-based supercomputer
to enter the TOP500 list, went into production
(\url{https://en.wikichip.org/wiki/supercomputers/astra}).
Furthermore, over 100 billion ARM processors have been produced as
of 2017 (\url{https://en.wikipedia.org/wiki/ARM_architecture}),
making it the most widely used instruction set architecture
in the world.

Clearly there are motivations for a large scientific computing
software library to support POWER and ARM architectures. We've extended
our continuous integration (CI) testing to include \texttt{ppc64le}
(POWER8 on Travis CI) and ARMv8 (on Shippable service). We also test
with the s390x architecture (IBM Z CPUs on Travis CI) so that we
can probe the behavior of our library on a big-endian machine.
This satisfies one of the major components of
improved CI testing laid out in a version of our
roadmap---specifically, ``CI for more exotic
platforms."

PEP 599 (\url{https://www.python.org/dev/peps/pep-0599/})
lays out a plan for new Python binary wheel
distribution support, \texttt{manylinux2014}, that adds
support for a number of architectures supported by the CentOS
Alternative Architecture Special Interest Group, including
ARMv8, ppc64le, as well as s390x. We are thus well-positioned
for a future where provision of binaries on these architectures
will be expected for a library at the base of the ecosystem.
